---
title: "Load counts, add metadata"
author: "Yuvarani Masarapu"
date: "7/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=16, fig.height=12, warning=FALSE, message=FALSE)
```

# R markdown script to load counts into a seurat object, and
# add metadata columns to the object.

### Loading package libraries
```{r}
library(Seurat)
library(SeuratObject)
library(readr)
library(readxl)
```

### Load processed data (from spaceranger) into seurat
```{r warning=FALSE,eval=FALSE}
#path to the directory where count matrices for each sample reside
data_dir <- "covidlung_counts"

data.list <- list.files(data_dir, recursive = F, full.names = T)

#iterate through all folders in data.list to load count matrices and images
data.object <- lapply(data.list, function(folder){
  each.object <- Load10X_Spatial(data.dir = folder)
})

rm(data_dir, data.list, dir)
```

### Add metadata

```{r}
lung_metadata <- read_excel("covidlung_metadata.xlsx")
```

```{r}
for (data in 1:length(data.object)){
  
  #create metadata columns for each sample object
  sample_type = c(rep(lung_metadata$sample_type[data],length(rownames(data.object[[data]]@meta.data))))
  tissue = c(rep(lung_metadata$tissue[data],length(rownames(data.object[[data]]@meta.data))))
  sub_array = c(rep(lung_metadata$sub_array[data],length(rownames(data.object[[data]]@meta.data))))
  viral_probes = c(rep(lung_metadata$viral_probes[data],length(rownames(data.object[[data]]@meta.data))))
  slide_id = c(rep(lung_metadata$slide_id[data], length(rownames(data.object[[data]]@meta.data))))
  sample_id = c(rep(lung_metadata$sample_id[data], length(rownames(data.object[[data]]@meta.data))))
  
  #convert to named vectors
  names(sample_type) <- rownames(data.object[[data]]@meta.data)
  names(tissue) <- rownames(data.object[[data]]@meta.data)
  names(sub_array) <- rownames(data.object[[data]]@meta.data)
  names(viral_probes) <- rownames(data.object[[data]]@meta.data)
  names(slide_id) <- rownames(data.object[[data]]@meta.data)
  names(sample_id) <- rownames(data.object[[data]]@meta.data)
  
  #add metadata columns to corresponding sample object
  data.object[[data]] <- AddMetaData(object = data.object[[data]], metadata = sample_type, col.name = "sample_type")
  data.object[[data]] <- AddMetaData(object = data.object[[data]], metadata = tissue, col.name = "tissue")
  data.object[[data]] <- AddMetaData(object = data.object[[data]], metadata = sub_array, col.name = "sub_array")
  data.object[[data]] <- AddMetaData(object = data.object[[data]], metadata = viral_probes, col.name = "viral_probes")
  data.object[[data]] <- AddMetaData(object = data.object[[data]], metadata = slide_id, col.name = "slide_id")
  data.object[[data]] <- AddMetaData(object = data.object[[data]], metadata = sample_id, col.name = "sample_id")
  
  rm(sample_type, tissue, sub_array, viral_probes, location, slide_id, sample_id)
}

rm(data)
```

### Subset the dataset for covid data only
```{r}
covid_genes_new <- read_csv("covid_genes.csv")
```

```{r}
data.covid = list()
for(index in 1:length(data.object)){
  data.covid[[index]] = data.object[[index]][(rownames(data.object[[index]]) %in% covid_genes_new$Covid_Gene), ]
}

rm(index)
```

### Add spot-level labels to to mark active viral signal captured from ST

  * Add metadata column to the data with spot labels:
    * "covid_spot" where covid gene is captured and "no_covid_spot" for all other spots
```{r}
for(index in 1:length(data.object)){
  
  #if it is a covid sample
  if(unique(data.object[[index]]$sample_type) %in% "covid"){
    
    #check if the sample has viral probes added
    if(unique(data.object[[index]]$viral_probes) %in% "yes")
    {
      barcodes = c(colnames(data.object[[index]]))
    
      #covid spots with atleast 1 UMI
      spots = colnames(data.covid[[index]])[data.covid[[index]]$nCount_Spatial > 0]
      for(i in 1:length(barcodes))
      {
        if(barcodes[i] %in% spots)
          {
          barcodes[i] <- "covid_spot"
          }
        else{
          barcodes[i] <- "no_covid_spot"}
      }
      names(barcodes) <- rownames(data.object[[index]]@meta.data)
      data.object[[index]] <- AddMetaData(object = data.object[[index]], metadata = barcodes, col.name = "spot_label")
    }
    
    #if the covid sample does not have a viral probes
    else{
      barcodes = c(colnames(data.object[[index]]))
      for(i in 1:length(barcodes)){
        barcodes[i] <- "no_covid_spot"
      }
      names(barcodes) <- rownames(data.object[[index]]@meta.data)
      data.object[[index]] <- AddMetaData(object = data.object[[index]], metadata = barcodes, col.name = "spot_label")
    }
  }
  
  #if the sample is a negavtive-control
  else{
    barcodes = c(colnames(data.object[[index]]))
    spots = colnames(data.covid[[index]])[data.covid[[index]]$nCount_Spatial > 0]
    for(i in 1:length(barcodes)){
      barcodes[i] <- "no_covid_spot"
    }
    names(barcodes) <- rownames(data.object[[index]]@meta.data)
    data.object[[index]] <- AddMetaData(object = data.object[[index]], metadata = barcodes, col.name = "spot_label")
  }
}

rm(index, barcodes, spots, i)
```

```{r}
saveRDS(data.object, file = "data.object.human.sarscov2.rds")

saveRDS(data.covid, file = "data.object.sarscov2.only.rds")
```

