---
title: "QC and filtering"
author: "Yuvarani Masarapu"
date: "7/29/2021"
output: html_document
---

### Loading package libraries
```{r}
library(SeuratObject)
library(Seurat)
library(ggplot2)
library(Matrix)
```


### Subset dataset to include only human geneset
```{r eval=FALSE}
data.object = readRDS(file = "data.object.human.sarscov2.rds")

for(index in 1:length(data.object)){
  data.object[[index]] = data.object[[index]][!(rownames(data.object[[index]]) %in% covid_genes_new$Covid_Gene), ]
}
```

### Quality Control

  * Filtering Hemoglobin genes

```{r}
hb.genes <- lapply(data.object, function(obj){
  p = grep(pattern = "^HB[GEMAQ]", x = rownames(obj), value = T)
  #delete the gene "HBEGF" which gets picked up but is not Hb gene
  p = p[-c(1)]
})

for(index in 1:length(data.object)){
  data.object[[index]] = data.object[[index]][!(rownames(data.object[[index]]) %in% hb.genes[[index]]), ]
}
```

#### UMI and Features distribution across the tissue

* Before filtering
```{r}
lapply(data.object, function(eachObj){
  VlnPlot(eachObj, features = c("nFeature_Spatial", "nCount_Spatial"), ncol = 2, pt.size = 0.1)
})
```

* Number of genes detected per UMI for each spot in the data
```{r}
for(index in 1:length(data.object)){
  
  data.object[[index]]$log10GenesPerUMI<-log10(data.object[[index]]$nFeature_Spatial) / log10(data.object[[index]]$nCount_Spatial) 
}
```

### Filtering for low quality spots

* Spot-level filtering: keeping spots with, 
  * nGene per spot >= 400
  * nUMI per spot >= 500
  * log10GenesPerUMI or genes/UMI ratio > 0.87

* Gene-level filtering
  * Keep gene if detected in atleast 1 spot

```{r eval=FALSE}
data.object.filter <- lapply(data.object, function(obj){
  obj = obj[rowSums(obj) > 0, 
            (obj$nCount_Spatial >= 500) & (obj$nFeature_Spatial >= 400) & (obj$log10GenesPerUMI > 0.87)]
})
```

* After filtering
```{r}
lapply(data.object.filter, function(obj){
   SpatialPlot(obj, features = c("nFeature_Spatial","nCount_Spatial"))
})
```

```{r}
saveRDS(data.object.filter, file = "data.object.filter.human.only.rds")
```


